---
title: "Homework 10"
author: "Brandon Simelgor"
date: "2023-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Brandon Simelgor
https://github.com/Simelgor05/Homework-10.git

```{r}
vehicle <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/vehicle.csv")
accident <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/accident.csv")
person <- read.csv("https://raw.githubusercontent.com/yumouqiu/DS202-Spring2023/main/Practice/data/fars2016/person.csv")
```


Part one: Accident data
are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
library(tidyverse)
head(accident)
accidentday <- accident %>%
  group_by(DAY_WEEK) %>%
  summarize(total_accidents = n())

ggplot(accidentday, aes(x = DAY_WEEK, y = total_accidents)) +
  geom_bar(stat = "identity") +
  xlab("Day") +
  ylab("Accidents") +
  ggtitle("Accidents by Day of the Week")
```

The days of the week where accidents happened more often were days 1, 6, and 7.

what time of the day do accidents happen (use variable HOUR)?
```{r}
accidentshour <- accident %>%
  filter(HOUR != 99) %>%
  group_by(HOUR) %>%
  summarize(total_accidents = n())

ggplot(accidentshour, aes(x = HOUR, y = total_accidents)) +
  geom_bar(stat = "identity") +
  xlab("Hour") +
  ylab("Accidents") +
  ggtitle("Accidents by Hour of the Day")
```

Accidents occured during every hour of the day but most frequently in th evening between the 15th and 24th hours of the day.

what is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?
```{r}
accidentsdrunkdriver <- accident %>%
  filter(DRUNK_DR > 0) %>%
  nrow()

accidentsdrunkdriver

```
There are 8720 accidents with at least one drunk driver.


Part two: Connecting data
Connect to the person table. Identify drivers (PER_TYP == 1, see fars manual ) and subset on them.
```{r}
head(person)
head(person) %>% filter(PER_TYP ==1)
```

Join accident and driver table (work out which variable(s) to use)
```{r}
head(vehicle)
accidentvehicle <- left_join(accident, vehicle, by = "ST_CASE")
```

Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!
```{r}
accidentsdayhourgender <- person %>%
  filter(HOUR != 99) %>%
  filter(SEX != 8) %>%
  filter(SEX != 9) %>%
  left_join(accidentvehicle, by = "ST_CASE") %>% 
  group_by(DAY_WEEK, HOUR, SEX) %>% 
  summarize(num_accidents = n())

accidentsdayhourgender

ggplot(accidentsdayhourgender, aes(x = HOUR, y = num_accidents, fill = DAY_WEEK)) +
  facet_wrap(~SEX)+
  geom_bar(stat = "identity", position = "stack") +
  xlab("Hour") +
  ylab("Accidents") +
  ggtitle("Accidents by Day of the Week, Hour of the Day, and Gender")
```

